Dominance Regions for Imaginary $\bfg$-vectors

### 2020-06-08 dominance_regions.tex
Lemma 3.3 proof started
Fixed statement
Stability of imaginary rays

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}

  Here we may take $\TT_2$ to have vertices labelled $t_\ell$ for $\ell\in\ZZ$, with $t_{2r}$ joined to $t_{2r+1}$ by an edge labelled $1$ and $t_{2r-1}$ joined to $t_{2r}$ by an edge labelled $2$.
  \begin{lemma}
    If $\lambda\in\ZZ^2\setminus\cI$, then $\cP(\lambda)=\{\lambda\}$.
  \end{lemma}
  \begin{proof}
    It is enough to consider $\phi_{t_{-2}}^{-1}\cC(\phi_{t_{-2}}\lambda) \cap \cC(\lambda) \cap \phi_{t_2}^{-1}\cC(\phi_{t_2}\lambda)$, we leave the details to the reader.
  \end{proof}

  \begin{lemma}
    \label{le:one direction}
    For $\lambda\in\cI\cap\ZZ^2$, 
    \begin{enumerate}
      \item $\cP_+(\lambda):=\bigcap_{i \ge 0}\phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)$ is the quadrilateral with corner vertices $\lambda$, $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$, and $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-b\lambda_2,\lambda_2)$.
      \item $\cP_-(\lambda):=\bigcap_{i \le 0}\phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)$ is the quadrilateral with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,-\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$, and $(\lambda_1,-c\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$.
    \end{enumerate}
  \end{lemma}

>>>>

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take $\TT_2$ to have vertices labelled $t_\ell$ for $\ell\in\ZZ$, with $t_{2r}$ joined to $t_{2r+1}$ by an edge labelled $1$ and $t_{2r-1}$ joined to $t_{2r}$ by an edge labelled $2$.

  Observe that $\phi_{t_2}=\phi_{t_{-2}}^{-1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} ((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  and $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} (-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

  By an eigenvector of a piecewise-linear map $\phi$ we mean a vector $\lambda$ such that for each $k\ge0$, there exists a scalar $\alpha_k$ so that $\phi^k(\lambda)=\alpha_k\lambda$.
  \begin{lemma}
    Any eigenvector of $\phi_{t_2}$ is a multiple of one of the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  \end{lemma}
  \begin{proof}
    First observe that the equation $\phi_{t_2}^k(\lambda)=\alpha_k\lambda$ cannot be satisfied for all $k\ge0$ unless $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$.
    In this region, $\phi_{t_2}$ is linear with eigenvalue $\alpha$ satisfying $\alpha^2-(bc-2)+1=0$, i.e. $\alpha=\frac{bc-2\pm\sqrt{bc(bc-4)}}{2}$.
    We thus require that 
    \[\frac{bc-2\pm\sqrt{bc(bc-4)}}{2}\lambda_1=(bc-1)\lambda_1+b\lambda_2 \qquad\text{and}\qquad \frac{bc-2\pm\sqrt{bc(bc-4)}}{2}\lambda_2= -c\lambda_1-\lambda_2,\]
    or
    \[\frac{-bc\pm\sqrt{bc(bc-4)}}{2b}\lambda_1=\lambda_2 \qquad\text{and}\qquad \frac{-bc\mp\sqrt{bc(bc-4)}}{2c}\lambda_2=\lambda_1.\]
    As these represent the same relation, the result follows.
  \end{proof}

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}

  \begin{lemma}
    If $\lambda\in\ZZ^2\setminus\cI$, then $\cP(\lambda)=\{\lambda\}$.
  \end{lemma}
  \begin{proof}
    It is enough to consider $\phi_{t_{-2}}^{-1}\cC(\phi_{t_{-2}}\lambda) \cap \cC(\lambda) \cap \phi_{t_2}^{-1}\cC(\phi_{t_2}\lambda)$, we leave the details to the reader.
  \end{proof}

  \begin{lemma}
    \label{le:one direction}
    For $\lambda\in\cI\cap\ZZ^2$, 
    \begin{enumerate}
      \item $\cP_+(\lambda):=\bigcap_{i \ge 0}\phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)$ is the quadrilateral with corner vertices $\lambda$, $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\lambda_2)$, and $(\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_1-b\lambda_2,\lambda_2)$.
      \item $\cP_-(\lambda):=\bigcap_{i \le 0}\phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)$ is the quadrilateral with corner vertices $\lambda$, $(\lambda_1+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_2,0)$, $(0,-\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$, and $(\lambda_1,-c\lambda_1+\frac{2-bc-\sqrt{bc(bc-4)}}{2}\lambda_2)$.
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    We prove (1) as (2) is obtained by interchanging $b$ with $c$ and swapping all ordered pairs.  
    We first show that $\cC(\lambda) \cap \phi_{t_i}^{-1}\cC(\phi_{t_i}\lambda)\supsetneq \cC(\lambda) \cap \phi_{t_j}^{-1}\cC(\phi_{t_j}\lambda)$ for $0\le i<j$.
  \end{proof}
